---
title: "script_statistics"
author: "Katja Kozjek"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare the environment

```{r load libraries, message=F, warning=F, echo=F}

library(dada2); packageVersion("dada2") 
library(phyloseq); packageVersion("phyloseq") 
library(vegan); packageVersion("vegan") 
library(ggplot2); packageVersion("ggplot2")
library(dplyr); packageVersion("dplyr")
library(tidyverse); packageVersion("tidyverse")
library(ggpubr); packageVersion("ggpubr")
library(devtools); packageVersion("devtools")
library(ape); packageVersion("ape")
library(picante); packageVersion("picante")
library(indicspecies); packageVersion("indicspecies")
library(ggvegan); packageVersion("ggvegan")
library(ggrepel); packageVersion("ggrepel")
library(ranacapa); packageVersion("ranacapa")

```

# Create phyloseq object, based on the UNITE taxonomy

```{r create phyloseq, echo=F}

asv_tab <- read.csv("data/csv/ASVs_counts.csv", sep = ",", row.names = 1)

#OTU table
OTU = otu_table(asv_tab, taxa_are_rows = FALSE)
OTUr = transform_sample_counts(OTU, function(x) x/sum(x))

#TAX table from UNITE
taxa_unite <- read.csv("data/csv/tax_unite2020.csv", sep = ",", row.names = 1)
tax_unite_table = tax_table(as.matrix(taxa_unite))
TAX = tax_table(tax_unite_table)

#META file 
meta_file = read.csv("data/csv/meta_DOK.csv", sep=",", row.names=1) 
meta_file = sample_data(data.frame(meta_file))

#create phyloseq
physeq_unite = phyloseq(OTU, TAX, meta_file)
physeq_unite

#phyloseq with relative abundance
physeq_unite_rel = phyloseq(OTUr, TAX, meta_file)

#save unite phyloseq object with relative abundance
saveRDS(physeq_unite_rel, "data/rds/phyloseq_unite_rel.rds")

#add random tree
unite_tree = rtree(ntaxa(physeq_unite), rooted=TRUE, tip.label=taxa_names(physeq_unite))
plot((unite_tree))

#add unite tree to phyloseq object 
physeq_unite = merge_phyloseq(physeq_unite, unite_tree)
physeq_unite

#save unite phyloseq object
saveRDS(physeq_unite, "data/rds/phyloseq_unite.rds")

```

## Filter phyloseq object
  
```{r filter phyloseq object, echo=F}

#remove non-AMF 
physeq_unite_Filtered <- subset_taxa(physeq_unite, !Phylum %in% c ("p__Basidiomycota", "p__Ascomycota"))
physeq_unite_Filtered

physeq_unite_Filtered_rel <- subset_taxa(physeq_unite_rel, !Phylum %in% c ("p__Basidiomycota", "p__Ascomycota"))

plot_bar(tax_glom(physeq_unite_Filtered_rel, "Phylum"), fill="Phylum")

sum(sample_sums(physeq_unite)) 
sum(sample_sums(physeq_unite_Filtered)) 
sort(sample_sums(physeq_unite_Filtered), decreasing = F) #we will discard samples with less than 500 reads 

```

## Phyloseq object with only AMF (Glomeromycota)

```{r AMF phyloseq object, echo=F}

#create new phyloseq 
physeq_AMF <- physeq_unite_Filtered

#save phyloseq object
saveRDS(physeq_AMF, "data/rds/phyloseq_AMF.rds")

write.csv(physeq_AMF@otu_table, "data/csv/AMF_count.csv")
write.csv(physeq_AMF@tax_table, "data/csv/tax_AMF.csv")

#compute prevalence of each feature, store as data.frame
prevdf=apply(X=otu_table(physeq_AMF), 
             MARGIN=ifelse(taxa_are_rows(physeq_AMF), yes=1, no=2),
             FUN=function(x){sum(x>0)})

#add taxonomy and total read frames to this data.frame
prevdf=data.frame(Prevalence=prevdf,
                  TotalAbundance=taxa_sums(physeq_AMF), tax_table(physeq_AMF))

#relative prevalence
#compute prevalence of each feature, store as data.frame
prevdf_rel=apply(X=otu_table(physeq_unite_Filtered_rel), 
                 MARGIN=ifelse(taxa_are_rows(physeq_unite_Filtered_rel), yes=1, no=2),
                 FUN=function(x){sum(x>0)})

#add taxonomy and total read frames to this data.frame
prevdf_rel=data.frame(Prevalence=prevdf_rel,
                      TotalAbundance=taxa_sums(physeq_unite_Filtered_rel), tax_table(physeq_unite_Filtered_rel))

write.csv(prevdf_rel,"results/prevalance_rel.csv")

```

# Rarefaction curves
  
```{r rarefaction curves, echo=F}

barplot(sort(sample_sums(physeq_AMF)),
        ylab= "Number of reads", xlab="Sample ID")

rare_curve <- rare_curve + xlab("Sequence sample size") + ylab("Number of ASVs") + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) 
print(rare_curve)

```
  
# Convert from phyloseq to vegan or from vegan to phyloseq
  
```{r vegan_phyloseq functions, echo=F}

# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}

```

# Prepare ASV level 
+ remove samples with less than 500 sequences
+ TSS normalization 

```{r asv level, echo=F}

sort(sample_sums(physeq_AMF), decreasing = F)

physeq_unite_Filtered_ASV <- prune_samples(sample_sums(physeq_unite_Filtered) >=500, physeq_AMF)

plot_bar((physeq_unite_Filtered_ASV))
unique(ntaxa(physeq_unite_Filtered_ASV)) 

#TSS normalization
unite_ASV_rel <- transform_sample_counts(physeq_unite_Filtered_ASV, function(x) x/sum(x))

```

# COMMUNITY COMPOSITION 
+ here I explore the community composition on genus level
+ agglomerate on genus level
+ PERMANOVA with Bray-Curtis
+ plot PCoA/NMDS

## GENUS level 
+ agglomerate on genus level 
+ remove samples with less than 500 reads
+ TSS normalization

```{r community composition genus level, echo=F}

## agglomerate on genus level 

length(get_taxa_unique(physeq_AMF, taxonomic.rank = "Genus"))

# agglomerate taxa at the genus level (combine all with the same name) and remove all taxa without genus level assignment

physeq_AMF_genus = tax_glom(physeq_AMF, "Genus", NArm = T)
physeq_AMF_genus #agglomerated on the genus level

sum(sample_sums(physeq_AMF_genus)) 
unique(tax_table(physeq_AMF_genus)[,6]) 
sort(sample_sums(physeq_AMF_genus), decreasing = F) #we will remove 3 samples

#keep samples with more than 500 reads 
physeq_AMF_genus_500 <- prune_samples(sample_sums(physeq_AMF_genus) >=500, physeq_AMF_genus)

sort(sample_sums(physeq_AMF_genus_500))
plot_bar((physeq_AMF_genus_500))
unique(ntaxa(physeq_AMF_genus_500))
any(taxa_sums(physeq_AMF_genus_500) == 0)

#TSS normalization 
physeq_AMF_genus_500rel <- transform_sample_counts(physeq_AMF_genus_500, function(x) x/sum(x))

``` 

# NMDS ordination (genus level)

```{r nmds ordination, echo=F}

#transfer OTU table from phyloseq into a vegan object
vegan_AMF_genus<-psotu2veg(physeq_AMF_genus_500rel)

#transfer sample data from phyloseq into a vegan object
veganSample_AMF_genus<-pssd2veg(physeq_AMF_genus_500rel)

#rename columns from ASV to genus name 
colnames(vegan_AMF_genus) <- c("g__Paraglomus",  "g__Acaulospora", "g__Septoglomus", "g__Glomus", "g__Claroideoglomus", "g__Gigaspora", "g__Funneliformis", "g__Diversispora", "g__Archaeospora", "g__Ambispora", "g__Dominikia",  "g__Palaeospora")
vegan_AMF_genus

#calculate the distance matrix
set.seed(0902)
nmds_ord <- metaMDS(vegan_AMF_genus, distance = "bray", k=2, trymax=1000, autotransform=T, permutations = 999)
nmds_ord$stress
nmds_ord$species

plot_ordination(physeq_AMF_genus_500rel, nmds_ord, color = "farming_system", shape = "time") + geom_point(size = 3) + theme_bw()

#stress is goodness of fit 
goodness(nmds_ord)
stressplot(nmds_ord)

#this function draws NMDS ordination diagram with sites (=my samples)
plot(nmds_ord)
plot(nmds_ord, display = 'sites', type = 't')
plot(nmds_ord, display = 'species', type = 't')

#species spores 
envfit(nmds_ord, vegan_AMF_genus) 

plot(nmds_ord, type='n')
points(nmds_ord, display=c('sites'), choices=c(1, 2), pch=3, col='red')
text(nmds_ord, display=c('species'), choices=c(1, 2), col='blue', cex=0.7)

```

```{r add VWC, echo=F}

#add volumetric soil water content (VWC)
veganSample_AMF_genus$VWC<-veganSample_AMF_genus$WC*veganSample_AMF_genus$BD
summary(veganSample_AMF_genus$VWC)

#rename
names(veganSample_AMF_genus)[names(veganSample_AMF_genus) == "SOC"] <- "total.C"

```

## prepare for envfit 

```{r calculate envfit, echo=F}

#envfit with metadata 
nmds_meta <- as(sample_data(physeq_AMF_genus_500rel), "data.frame")
env_nmds <- nmds_meta[,5:24]

data.scores_nmds = as.data.frame(scores(nmds_ord, display = 'sites')) #save NMDS to dataframe
species.scores_nmds = as.data.frame(scores(nmds_ord, display = 'species'))

#add grouping variable to dataframe
data.scores_nmds <- cbind(as.data.frame(data.scores_nmds), farming_system = env_nmds$farming_system, 
                          treatment = env_nmds$treatment)
head(data.scores_nmds)

```

## calculate envfit for ndms ordination
+ investigate the soil parameters which may be driving the distribution pattern

```{r envfit for nmds, echo=F}

#calculate envfit
set.seed(0935)
env.fit <- envfit(scores(nmds_ord), env_nmds, permutations = 999, na.rm = TRUE, strata=nmds_meta$block)
env.fit

#extract scores from environment
env.scores <- as.data.frame(scores(env.fit, display = "vectors"))
env.scores <- cbind(env.scores, env.variables = rownames(env.scores), pval = env.fit$vectors$pvals)
head(env.scores)
sig.env <- subset(env.scores, pval<=0.05) #extract only significant soil parameters
head(sig.env)

#envfit write table
envfit_nmds <- data.frame((env.fit$vectors)$r, (env.fit$vectors)$pvals)

#vectors as continuous, factors as categorical
#vectors are soil parameters and factors (centroids) are treatment + farming system
en_coord_cont = as.data.frame(scores(env.fit, "vectors")) * ordiArrowMul(env.fit)
en_coord_cat = as.data.frame(scores(env.fit, "factors")) * ordiArrowMul(env.fit)

```

# PERMANOVA with BC on genus level 

```{r permanova on genus level, echo=F}

unite_genus_rel <- transform_sample_counts(physeq_AMF_genus_500rel, function(x) x/sum(x))
bray_unite_genus <- phyloseq::distance(unite_genus_rel, method = "bray")

meta.unite_genus <- as(sample_data(physeq_AMF_genus_500rel), "data.frame")

set.seed(1546)
perm.genus <- adonis2(bray_unite_genus ~ treatment+time+farming_system+
                        farming_system:treatment+
                        treatment:time+
                        time:farming_system,
                        strata=meta.unite_genus$block,
                        data = meta.unite_genus,permutations = 999)
perm.genus
```
  
#Multivariate dispersion 
+ As PERMANOVA is sensitive to unequal multivariate dispersion, multivariate group means were tested 

```{r multivariate dispersion, echo=F}

modSystem <- with(veganSample_AMF_genus,  betadisper(bray_unite_genus, farming_system))
modSystem

modDrought <- with(veganSample_AMF_genus,  betadisper(bray_unite_genus, treatment))
modDrought

modTime <- with(veganSample_AMF_genus,  betadisper(bray_unite_genus, time))
modTime

set.seed(1089)
permutest(modSystem)

set.seed(1089)
permutest(modDrought)

set.seed(1089)
permutest(modTime)

```

# Constrained analysis with db-RDA, to highlight effects
+ Plot only the variation in your data that explains changes in community composition

## calculate dbrda

```{r plot dbrda, echo=F}

#create the distance matrix
set.seed(1120)
dbrda <- dbrda(vegan_AMF_genus ~ farming_system+time+Condition(block),veganSample_AMF_genus,dist="bray")
head(scores(dbrda, display = 'sites'))
smry <- summary(dbrda)

sppscores(dbrda) <- vegan_AMF_genus

anova(dbrda)  # overall test of the significant of the analysis
anova(dbrda, by="axis")
anova(dbrda, by="term", permutations = 999)

ordiplot(dbrda,type="t",scaling=2)
ordiplot(dbrda,type="t",scaling=1)

#check constrained and unconstrained variance
dbrda$CCA$tot.chi/dbrda$tot.chi

constrained_eig <- dbrda$CCA$eig/dbrda$tot.chi*100
unconstrained_eig <- dbrda$CA$eig/dbrda$tot.chi*100
expl_var <- c(constrained_eig, unconstrained_eig)
barplot (expl_var[1:20], col = c(rep ('red', length (constrained_eig)), rep ('black', length (unconstrained_eig))),
         las = 2, ylab = '% variation')

#calculate amount of variation explained by the axis
b_xlabel<-paste0("db-RDA1"," ","[", round(((dbrda$CCA$eig[1]/dbrda$CA$tot)*100),1), "%]" )
b_ylabel<-paste0("db-RDA2"," ","[", round(((dbrda$CCA$eig[2]/dbrda$CA$tot)*100),1), "%]" )

#fortify: transfer into data frame that can be used for ggplot
scrs <- fortify(dbrda)

#select site (-> Samples)
sites <- subset(scrs, subset = Score == "sites") 

#select species (-> taxa)
species <- subset(scrs, subset = Score == "species")

(Time<-veganSample_AMF_genus$time) #choose group variable here
(System<-veganSample_AMF_genus$farming_system) #choose shape variable
df <- cbind(sites, Time , System) ## add on the group variable

df$Time <- factor(df$Time, levels=c("4w", "13w"), labels = c("4 weeks", "13 weeks"))

```

## plot original dbRDA

```{r dbrda original, echo=F}

dbrda_original <- ggplot(df, aes(x = dbRDA1, y = dbRDA2)) + geom_point(aes(colour = System, shape = Time), size = 5,alpha=0.8) +
scale_color_manual(values = c("dimgrey", "darkgray")) + labs(x = paste0("db-RDA1"," ","[", round(((dbrda$CCA$eig[1]/dbrda$CA$tot)*100),1), "%]" ), y = paste0("db-RDA2"," ","[", round(((dbrda$CCA$eig[2]/dbrda$CA$tot)*100),1), "%]" )) + 
  geom_vline(xintercept = 0, linetype="dotted") +
  geom_hline(yintercept = 0, linetype="dotted") +
  theme_bw() + 
  theme(text = element_text(size = 14)) + 
  theme(axis.title = element_text(size = 12, colour = "black"), axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), legend.title = element_text(size = 11, colour = "black"), legend.text = element_text(size = 10, colour = "black")) + labs(colour = "Farming system", shape= "Sampling time") + xlab("dbRDA1 (18.5% of fitted, 12.4% of total variation)") + ylab("dbRDA2 (7.7% of fitted, 5.3% of total variation)") 

print(dbrda_original)
```

## add environmental parameters to dbrda

```{r dbrda environment, echo=F}
## environmental variables 

names(veganSample_AMF_genus)
env_dbrda <- veganSample_AMF_genus[,2:25]
env_dbrda <- env_dbrda[,-c(2,3,4,7)]
names(env_dbrda)

#rename environmental parameters
names(env_dbrda)[names(env_dbrda) == "total.P"] <- "total P"
names(env_dbrda)[names(env_dbrda) == "total.N"] <- "total N"
names(env_dbrda)[names(env_dbrda) == "total.C"] <- "total C"
names(env_dbrda)[names(env_dbrda) == "total.C"] <- "total C"
names(env_dbrda)[names(env_dbrda) == "C.N"] <- "C:N ratio"
names(env_dbrda)[names(env_dbrda) == "NLFA"] <- "AMF biomass"
names(env_dbrda)[names(env_dbrda) == "N.roots"] <- "N roots"
names(env_dbrda)[names(env_dbrda) == "C.roots"] <- "C roots"
names(env_dbrda)[names(env_dbrda) == "crop.biomass"] <- "crop biomass"
names(env_dbrda)[names(env_dbrda) == "weed.cover"] <- "weed cover"
names(env_dbrda)[names(env_dbrda) == "plant.height"] <- "plant height"
names(env_dbrda)[names(env_dbrda) == "N.shoots"] <- "N shoots"
names(env_dbrda)[names(env_dbrda) == "C.shoots"] <- "C shoots"
names(env_dbrda)[names(env_dbrda) == "root.DW"] <- "root DW"

#sites (=my samples)
data.scores_dbrda = as.data.frame(scores(dbrda, display = 'sites')) #save to dataframe

#extract species
species.scores_dbrda = as.data.frame(scores(dbrda, display = 'species'))

#add grouping variable to dataframe
data.scores_dbrda <- cbind(as.data.frame(data.scores_dbrda), farming_system = env_dbrda$farming_system, time = env_dbrda$time)
head(data.scores_dbrda)

#rename columns 
names(data.scores_dbrda)[names(data.scores_dbrda) == "farming_system"] <- "Farming system"
names(data.scores_dbrda)[names(data.scores_dbrda) == "time"] <- "Sampling time"

set.seed(1722)
envfit_dbrda <- envfit(dbrda, env_dbrda, strata=veganSample_AMF_genus$block,na.rm = TRUE, display="lc")
envfit_dbrda

#extract scores from environment
env.scores_dbrda <- as.data.frame(scores(envfit_dbrda, display = "vectors"))
env.scores_dbrda <- cbind(env.scores_dbrda, env.variables = rownames(env.scores_dbrda), pval = envfit_dbrda$vectors$pvals)

head(env.scores_dbrda)
sig.env_dbrda <- subset(env.scores_dbrda, pval<=0.05) #extract only significant soil parameters
View(sig.env_dbrda)

#vectors as continuous, factors as categorical
#vectors are soil parameters and factors (centroids) are time + farming system
en_coord_cont_dbrda = as.data.frame(scores(envfit_dbrda, "vectors")) * ordiArrowMul(envfit_dbrda)
en_coord_cat_dbrda = as.data.frame(scores(envfit_dbrda, "factors")) * ordiArrowMul(envfit_dbrda)

#rename factors

rownames(en_coord_cat_dbrda)[rownames(en_coord_cat_dbrda) == "farming_systemBIODYN"] <- "BIODYN"
rownames(en_coord_cat_dbrda)[rownames(en_coord_cat_dbrda) == "farming_systemCONMIN"] <- "CONMIN"
rownames(en_coord_cat_dbrda)[rownames(en_coord_cat_dbrda) == "time4w"] <- "4 weeks"
rownames(en_coord_cat_dbrda)[rownames(en_coord_cat_dbrda) == "time13w"] <- "13 weeks"

#envfit for species
spp.fit <- envfit(dbrda, env_dbrda, strata=veganSample_AMF_genus$block,na.rm = TRUE, display="lc")

```

## plot original dbRDA with environmental parameters 

```{r plot dbRDA and environmental parameters, echo=F}

dbrda_original_env <- dbrda_original + geom_segment(data = sig.env_dbrda,
                                              aes(x = 0, xend = dbRDA1,
                                                  y = 0, yend = dbRDA2),
                                              arrow = arrow(length = unit(2.5, "mm")),
                                              colour = "black",
                                              alpha=1,
                                              size=0.6,
                                              lineend ="round",
                                              linejoin="round") + 
  ggrepel::geom_text_repel(data = sig.env_dbrda, aes(x=dbRDA1, y=dbRDA2, label =env.variables), cex = 4,                color="black", direction = "both", segment.size = 0.25) + geom_point(data = en_coord_cat_dbrda, aes(x                = dbRDA1, y = dbRDA2), shape = "diamond", size = 8, alpha = 0.8, colour = "black") + geom_text(data =                en_coord_cat_dbrda, aes(x = dbRDA1, y = dbRDA2), label = row.names(en_coord_cat_dbrda), colour =                     "black", fontface = "bold") + theme(panel.grid.major = element_blank(), panel.grid.minor =                           element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

print(dbrda_original_env)

```

# INDICATOR SPECIES ANALYSIS
+ to investigate if there are indicators of the farming system, the time of sampling and drought treatment

```{r indicator species analysis, echo=F}

#Transfer Phyloseq into a vegan ASV table
ind_analysis <- psotu2veg(physeq_AMF)

#meta file for indicators 
ind_meta <- as(sample_data(physeq_AMF), "data.frame")

ind_time <- ind_meta$time
ind_treatment <- ind_meta$treatment
ind_system <- ind_meta$farming_system

length(unique(ind_treatment))
length(unique(ind_time))
length(unique(ind_system))

#indicators for farming system 
set.seed(1809)
indval_System = multipatt(ind_analysis,ind_system,func = "IndVal", control = how(nperm=999))
summary(indval_System,indvalcomp=TRUE)

#indicators for drought treatment
set.seed(1809)
indval_Treatment = multipatt(ind_analysis,ind_treatment,func = "IndVal", control = how(nperm=999))
summary(indval_Treatment,indvalcomp=TRUE)

#indicators for the time of sampling
set.seed(1809)
indval_Time = multipatt(ind_analysis,ind_time,func = "IndVal", control = how(nperm=999))
summary(indval_Time,indvalcomp=TRUE)

```

```{r separate indicators for drought within farming system}

ind_analysis_biodyn <- ind_analysis[-c(1,2,3,7,8,15,16,17,21,22,23,24,25,26,30,31,32,38,39,40,44,45,46),]

ind_analysis_conmin <- ind_analysis[-c(4,5,6,9,10,11,12,13,14,18,19,20,27,28,29,33,34,35,36,37,41,42,43),]

#meta file for the indicators 
indicators_biodyn <- droplevels(subset(ind_meta, ind_meta$farming_system=="BIODYN"))
indicators_conmin <- droplevels(subset(ind_meta, ind_meta$farming_system=="CONMIN"))

biodyn_treatment <- indicators_biodyn$treatment
conmin_treatment <- indicators_conmin$treatment

biodyn_drought <- multipatt(ind_analysis_biodyn,biodyn_treatment,func = "IndVal", control = how(nperm=999))
summary(biodyn_drought,indvalcomp=TRUE)

conmin_drought <- multipatt(ind_analysis_conmin,conmin_treatment,func = "IndVal", control = how(nperm=999))
summary(conmin_drought,indvalcomp=TRUE) 

```

```{r separate indicators for drought within times}

ind_analysis_T1 <- ind_analysis[-c(24:46),]

ind_analysis_T3 <- ind_analysis[-c(1:23),]

#meta file for the indicators 
indicators_T1 <- droplevels(subset(ind_meta, ind_meta$time=="4w"))
indicators_T3 <- droplevels(subset(ind_meta, ind_meta$time=="13w"))

T1_treatment <- indicators_T1$treatment
T3_treatment <- indicators_T3$treatment

T1_drought <- multipatt(ind_analysis_T1,T1_treatment,func = "IndVal", control = how(nperm=999))
summary(T1_drought,indvalcomp=TRUE)

T3_drought <- multipatt(ind_analysis_T3,T3_treatment,func = "IndVal", control = how(nperm=999))
summary(T3_drought,indvalcomp=TRUE)

```

# TAXONOMIC COMPOSITION on different taxonomic levels
+ here we take a file with all ASVs with the relative abundances

## Class level
+ 3 classes 
+ Glomeromycetes as the most abundant 

```{r investigate class level, echo=F}

#taxonomy_AMF

unique(tax_table(unite_ASV_rel)[,3])

subset_class <- subset_taxa(unite_ASV_rel, Class %in% c("c__Paraglomeromycetes", "c__Glomeromycetes", "c__Archaeosporomycetes")) 

subset_class
subset_class = prune_samples(sample_sums(subset_class)>0, subset_class)

data_class <- subset_class %>%
  tax_glom(taxrank = "Class") %>%     #agglomerate at class level
  psmelt() %>%                        #melt to long format
  arrange(Class)                      #sort data frame alphabetically by class

# change the order of varible levels with factor()
data_class$treatment <- factor(data_class$treatment, levels = c("R", "RC", "C"))

#plot AMF class 
ggplot(data_class) + geom_col(mapping = aes(x = treatment:farming_system, y = Abundance, fill = Class), position = "fill", show.legend = TRUE) + theme(axis.text.x = element_text(angle = 35, hjust = 1)) + scale_x_discrete(name ="Treatment:Farming system")  + facet_wrap(~time) + scale_fill_discrete(name="AMF class", breaks=c("c__Archaeosporomycetes", "c__Glomeromycetes", "c__Paraglomeromycetes"), labels=c("Archaeosporomycetes", "Glomeromycetes", "Paraglomeromycetes")) + ylab("Proportion")

```

## All orders 
+ there are 5 orders

```{r all orders AMF, echo=F}

unique(tax_table(unite_ASV_rel)[,4])

subset_orders_AMF <- subset_taxa(unite_ASV_rel, Order %in% c("o__Diversisporales","o__Glomerales","o__Gigasporales", "o__Paraglomerales", "o__Archaeosporales"))

orders_AMF = prune_samples(sample_sums(subset_orders_AMF)>0, subset_orders_AMF)

orders_AMF <- orders_AMF %>%
  tax_glom(taxrank = "Order") %>%      #agglomerate at order level
  psmelt() %>%                         #melt to long format
  arrange(Order)


# change the order of varible levels with factor()
orders_AMF$treatment <- factor(orders_AMF$treatment, levels = c("R", "RC", "C"))

ggplot(orders_AMF) + geom_col(mapping = aes(x = treatment:farming_system, y = Abundance, fill = Order), position = "fill", show.legend = TRUE)  + theme(axis.text.x = element_text(angle = 35, hjust = 1)) + scale_x_discrete(name ="Treatment:Farming system") + scale_fill_discrete(name="AMF order", breaks=c("o__Archaeosporales", "o__Diversisporales", "o__Glomerales", "o__Paraglomerales"), labels=c("Archaeosporales", "Diversisporales", "Glomerales", "Paraglomerales")) +ylab("Proportion")

```

## Focus on Glomeromycetes class
+ focus on Glomeromyctes class, as it is the most abundant one and three orders within it

```{r Glomeromycetes class, echo=F}

unique(tax_table(unite_ASV_rel)[,4])

subset_orders <- subset_taxa(unite_ASV_rel, Order %in% c("o__Diversisporales","o__Glomerales","o__Gigasporales"))

glomeromycetes_orders = prune_samples(sample_sums(subset_orders)>0, subset_orders)

glomeromycetes_orders <- glomeromycetes_orders %>%
  tax_glom(taxrank = "Order") %>%      #agglomerate at order level
  psmelt() %>%                         #melt to long format
  arrange(Order)                       #sort data frame alphabetically by order

# change the order of varible levels with factor()
glomeromycetes_orders$treatment <- factor(glomeromycetes_orders$treatment, levels = c("R", "RC", "C"))

ggplot(glomeromycetes_orders) + geom_col(mapping = aes(x = treatment:farming_system, y = Abundance, fill = Order), position = "fill", show.legend = TRUE) + theme(axis.text.x = element_text(angle = 35, hjust = 1)) + scale_fill_discrete(name="Glomeromycetes class", breaks=c("o__Diversisporales", "o__Gigasporales", "o__Glomerales"), labels=c("Diversisporales", "Gigasporales", "Glomerales")) + facet_wrap(~time) + ylab("Proportion")

```

## Family level 
+ there are 8 families 

```{r family level, echo=F}

unique(tax_table(unite_ASV_rel)[,5])

subset_family <- subset_taxa(unite_ASV_rel, Family %in% c("f__Archaeosporaceae", "f__Glomeraceae", "f__Diversisporaceae", "f__Ambisporaceae", "f__Acaulosporaceae", "f__Paraglomeraceae", "f__Gigasporaceae", "f__Claroideoglomeraceae")) 

subset_family
subset_family = prune_samples(sample_sums(subset_family)>0, subset_family)

data_family <- subset_family %>%
  tax_glom(taxrank = "Family") %>%
  psmelt() %>%
  arrange(Family)

# change the order of varible levels with factor()
data_family$treatment <- factor(data_family$treatment, levels = c("R", "RC", "C"))

ggplot(data_family) + geom_col(mapping = aes(x = treatment:farming_system, y = Abundance, fill = Family), position = "fill", show.legend = TRUE) + theme(axis.text.x = element_text(angle = 35, hjust = 1)) + scale_x_discrete(name ="Treatment:Farming system") + scale_fill_discrete(name="AMF family", breaks=c("f__Acaulosporaceae", "f__Ambisporaceae", "f__Archaeosporaceae", "f__Claroideoglomeraceae", "f__Diversisporaceae", "f__Gigasporaceae","f__Glomeraceae",  "f__Paraglomeraceae"), labels=c("Acaulosporaceae", "Ambisporaceae", "Archaeosporaceae", "Claroideoglomeraceae", "Diversisporaceae", "Gigasporaceae", "Glomeraceae", "Paraglomeraceae")) + facet_wrap(~time) + ylab("Proportion")

```

## Genus level 
+ 12 genera
+ here we include NA

```{r genus level, echo=F}

#here we will include unassigned

unique(tax_table(unite_ASV_rel)[,6])

subset_genera <- subset_taxa(unite_ASV_rel, Genus %in% c("g__Acaulospora", "g__Glomus", "g__Palaeospora", "NA", "g__Diversispora", "g__Funneliformis", "g__Paraglomus", "g__Ambispora", "g__Claroideoglomus", "g__Gigaspora", "g__Dominikia", "g__Archaeospora", "g__Septoglomus"))

subset_genera = prune_samples(sample_sums(subset_genera)>0, subset_genera)

data_genera <- subset_genera %>%
  tax_glom(taxrank = "Genus") %>%
  psmelt() %>%
  arrange (Genus)

# change the order of varible levels with factor()
data_genera$treatment <- factor(data_genera$treatment, levels = c("R", "RC", "C"))
data_genera$time <- factor(data_genera$time, levels=c("4w", "13w"), labels = c("4 weeks", "13 weeks"))

genusAMF_abundance <- ggplot(data_genera) + geom_col(mapping = aes(x = treatment, y = Abundance, fill = Genus), position = "fill", show.legend = TRUE) + scale_x_discrete(name ="Drought treatment") + scale_fill_manual(values = c("#EAD3BF", "#AA9486", "#B6854D", "#7294D4","#F1BB7B", "#FD6467", "#9C964A", "#CDC08C", "#FAD77B","#F3DF6C", "#CEAB07","#D3DDDC"), name="AMF genera", breaks=c("g__Acaulospora", "g__Ambispora", "g__Archaeospora", "g__Claroideoglomus", "g__Diversispora", "g__Dominikia", "g__Funneliformis", "g__Gigaspora", "g__Glomus", "g__Palaeospora", "g__Paraglomus", "g__Septoglomus"), labels=c("Acaulospora", "Ambispora", "Archaeospora", "Claroideoglomus", "Diversispora", "Dominikia", "Funneliformis", "Gigaspora", "Glomus", "Palaeospora", "Paraglomus", "Septoglomus")) + ylab("Proportion of relative abundance") + theme_bw() + facet_wrap(time~farming_system)

print(genusAMF_abundance)

```

# Session Info

```{r sessionInfo, include=TRUE, echo=TRUE}
sessionInfo()
```